\ourlongtitle

### 2021-08-27 Bernoulli/AnswerReviews.tex

We would like to thank the associate editor and the referees for the attention they paid to our work and for their positive feedbacks.

\bigskip

Minor errors and typos listed by R2 were corrected as suggested. Please find below a few detailed answers.

\begin{itemize}
\item It might have made the paper less scary and dry to read to first provide standard setups (along the lines of what is at the top of p.2), then proceed to the objective (formula (1.2)) and only then supplement all the measure-theoretic setup, e.g. as part of section 2. This would also have yielded the advantage of collecting mathematical notation at the start of section 2 (Is a collection of common notation in tabular form allowed in Bernoulli? This would be a friendly service to the reader.) rather than scattering it over the start of section 1 and section 2.\\
{\em We thank the referee for this suggestion. We postponed the introduction of (1.2)  and the objective of the paper (additive smoothing) after the standard setups (Feynman-Kac path models and hidden Markov models).}
\item p.7 (2.10) and elsewhere uses rectangular brackets as in $\phi_m[\ell_m(\cdot,x_{m+1})]$ which I think is meant to denote $\int_{X_m}\ell_m(x_m,x_{m+1})\phi_m(dx_m)$. This notation has not been introduced. The "Preliminaries" at the top of p.4 have $\mu h$ for the integral of a function $h$ against a measure $\mu$ which would suggest $\phi_m \ell(\cdot,x_{m+1})$. Since the rectangular brackets are used in this sense at many points in the paper, perhaps this notation should be added to the preliminaries.\\
{\em }
\item p.7 Just before 2.2.2, it would be nice to comment on where one might get suitable adjustment weights $\vartheta_n$ from. How are these to be chosen or tuned?\\
{\em To do!}
\item p. 15-16: for the presentation of Example 2, I feel it would be better to describe the problem fully before giving Durham and Gallant's solution.\\
{\em Do not agree!}
\item p.29 Appendix A in the supplement has $\phi_n(dx_n)[\ell(\cdot,x_{n+1})]$  in the numerator that arises from $B_n(x_{n+1},dx_n)$. I think this should read $\phi_n(dx_n)\ell(x_n,x_{n+1})$  instead?\\
{\em The normalizing term given in the appendix is correct, see (2.10).}
\item p.30 Appendix A has $\phi_{0:n}L_{0,n}1_{X^n}$  in the denominator for  $\phi_{0:n+1}h$ which I think should read  $\phi_{0:n}L_{0,n}1_{X^{n+1}}$ as there is a dangling $dx_{n+1}$ otherwise.\\
{\em There was indeed a mistake in this identity, thank you for pointing this out. This was corrected using (2.3).}
\end{itemize}

### 2021-08-27 Bernoulli/glo-2020-intro.tex
Answer + a few modifications in the introduction

Let $(\Xset_n, \Xfd_n)_{n \in \nset}$ be a sequence of general state spaces and let, for all $n \in \nset$, $\uk{n} : \Xset_n \times \Xfd_{n + 1} \to \rsetnn$ be bounded kernels in the sense that $\sup_{x \in \Xset_n} \uk{n}(x, \Xset_{n + 1}) < \infty$. We will assume a dominated model where each kernel $\uk{n}$ has a kernel density $\ud{n}$ with respect to some $\sigma$-finite reference measure $\mu_{n + 1}$ on $\Xfd_{n + 1}$. Finally, let $\chi$ be some bounded measure on $\Xfd_0$. In the following, we denote state-space product sets and $\sigma$-fields by $\Xset^n \eqdef \Xset_0 \times \cdots \times \Xset_n$ and $\Xfd^n \eqdef \Xfd_0 \tensprod \cdots \tensprod \Xfd_n$, respectively, and consider probability measures  
\begin{equation} \label{eq:def:post}
\post{0:n}(\rmd x_{0:n}) \propto \chi(\rmd x_0) \prod_{m = 0}^{n - 1} \uk{m}(x_m, \rmd x_{m + 1}), \quad n \in \nset, 
\end{equation}
on these product spaces.\footnote{We will always use the standard convention $\prod_{\varnothing} = 1$, implying that $\post{0} \propto \chi$.} Given a sequence $(\addf{n})_{n \in \nset}$ of measurable functions $\addf{n} : \Xset_n \times \Xset_{n + 1} \to \rset$, the aim of the present paper is the online approximation of expectations of \emph{additive functionals}   
\begin{equation} \label{eq:add:func}
    h_n : \Xset^n \ni x_{0:n} \mapsto \sum_{m = 0}^{n - 1} \addf{m}(x_m, x_{m + 1})
\end{equation}
under the distribution flow $(\post{0:n})_{n \in \nset}$ using \emph{sequential Monte Carlo} (SMC) methods.

The generality of the model \eqref{eq:def:post} is striking. In the special case where each $\uk{n}$ can be decomposed as $\uk{n}(x_n, \rmd x_{n + 1}) = \md{n}(x_n) \, \hk_n(x_n, \rmd x_{n + 1})$ for some Markov kernel $\hk_n$ and some nonnegative potential function $\md{n}$, \eqref{eq:def:post} yields the \emph{Feynman-Kac path models} \cite{delmoral:2004}, which are applied in a large variety of scientific and engineering disciplines, including statistics, physics, biology, and signal processing. In a \emph{hidden Markov model} (HMM) (see, \eg,  \cite{Cappe:2005:IHM:1088883}), a Markov chain $(X_n)_{n \in \nset}$ with kernels $(\hk_n)_{n \in \nset}$ and initial distribution $\chi$ is only partially observed through a sequence $(Y_n)_{n \in \nset}$ of observations being conditionally independent given the Markov states. In that case, $g_n$ plays the role of the likelihood of the state $X_n$ given the corresponding observation $Y_n$, and $\post{0:n}$ describes the \emph{joint-smoothing distribution}, \ie, the joint posterior of the hidden states $X_0, \ldots, X_n$ given corresponding observations (see Example~\ref{ex:state-space:models} for details). We will adopt this terminology throughout the present paper and refer to the distributions defined in \eqref{eq:def:post} as `smoothing distributions'; the problem of computing expectations of functionals of type \eqref{eq:add:func} under these distributions will be referred to as `additive smoothing'. General state-space HMMs are prevalent in time-series and sequential-data analysis and are used extensively in, \eg, movement ecology \cite{michelot2016movehmm}, energy-consumption modeling \cite{candanedo2017methodology}, genomics \cite{yau2011bayesian}, target tracking \cite{sarkka2007rao}, enhancement and segmentation of speech and audio signals \cite{rabiner1989tutorial}; see also \cite{sarkka2013bayesian, zucchini2017hidden} and the numerous references therein. Operating on models of this sort, online additive smoothing is instrumental for, \emph{e.g.},    
\begin{enumerate}
    \item[--] \emph{path reconstruction}, \ie, the estimation of hidden states given observations. Especially in \emph{fixed-point smoothing}, where interest is in computing the expectations of $h(X_m)$ conditionally on $Y_0, \ldots, Y_n$ for some given $m$ and test function $h$ as $n$ tends to infinity, a problem that can be cast into our framework by letting, in \eqref{eq:add:func}, $\addf{m}(x_m, x_{m + 1}) = h(x_m)$ and $\addf{\ell} \equiv 0$ for all $\ell \neq m$. 
    \item[--] \emph{parameter inference}, where additive smoothing is a key ingredient in the computation of log-likelihood gradients (\emph{score functions}) via Fisher's identity or the intermediate quantity of the \emph{expectation-maximisation} (EM) \emph{algorithm}; see, \eg, \cite[Chapter~10]{Cappe:2005:IHM:1088883}. On-the-fly computation becomes especially important in online implementations via, \eg, the \emph{online EM} or \emph{recursive maximum likelihood} approaches \cite{cappe:2009,legland:mevel:1997}.  
 \end{enumerate}

>>>>

Let $(\Xset_n, \Xfd_n)_{n \in \nset}$ be a sequence of general state spaces and let, for all $n \in \nset$, $\uk{n} : \Xset_n \times \Xfd_{n + 1} \to \rsetnn$ be bounded kernels in the sense that $\sup_{x \in \Xset_n} \uk{n}(x, \Xset_{n + 1}) < \infty$. We will assume a dominated model where each kernel $\uk{n}$ has a kernel density $\ud{n}$ with respect to some $\sigma$-finite reference measure $\mu_{n + 1}$ on $\Xfd_{n + 1}$. Finally, let $\chi$ be some bounded measure on $\Xfd_0$. In the following, we denote state-space product sets and $\sigma$-fields by $\Xset^n \eqdef \Xset_0 \times \cdots \times \Xset_n$ and $\Xfd^n \eqdef \Xfd_0 \tensprod \cdots \tensprod \Xfd_n$, respectively, and consider probability measures  
\begin{equation} \label{eq:def:post}
\post{0:n}(\rmd x_{0:n}) \propto \chi(\rmd x_0) \prod_{m = 0}^{n - 1} \uk{m}(x_m, \rmd x_{m + 1}), \quad n \in \nset, 
\end{equation}
on these product spaces.\footnote{We will always use the standard convention $\prod_{\varnothing} = 1$, implying that $\post{0} \propto \chi$.}

The generality of the model \eqref{eq:def:post} is striking. In the special case where each $\uk{n}$ can be decomposed as $\uk{n}(x_n, \rmd x_{n + 1}) = \md{n}(x_n) \, \hk_n(x_n, \rmd x_{n + 1})$ for some Markov kernel $\hk_n$ and some nonnegative potential function $\md{n}$, \eqref{eq:def:post} yields the \emph{Feynman-Kac path models} \cite{delmoral:2004}, which are applied in a large variety of scientific and engineering disciplines, including statistics, physics, biology, and signal processing. In a \emph{hidden Markov model} (HMM) (see, \eg,  \cite{Cappe:2005:IHM:1088883}), a Markov chain $(X_n)_{n \in \nset}$ with kernels $(\hk_n)_{n \in \nset}$ and initial distribution $\chi$ is only partially observed through a sequence $(Y_n)_{n \in \nset}$ of observations being conditionally independent given the Markov states. In that case, $g_n$ plays the role of the likelihood of the state $X_n$ given the corresponding observation $Y_n$, and $\post{0:n}$ describes the \emph{joint-smoothing distribution}, \ie, the joint posterior of the hidden states $X_0, \ldots, X_n$ given corresponding observations (see Example~\ref{ex:state-space:models} for details). We will adopt this terminology throughout the present paper and refer to the distributions defined in \eqref{eq:def:post} as `smoothing distributions'.

Given a sequence $(\addf{n})_{n \in \nset}$ of measurable functions $\addf{n} : \Xset_n \times \Xset_{n + 1} \to \rset$, the aim of the present paper is the online approximation of expectations of \emph{additive functionals}   
\begin{equation} \label{eq:add:func}
    h_n : \Xset^n \ni x_{0:n} \mapsto \sum_{m = 0}^{n - 1} \addf{m}(x_m, x_{m + 1})
\end{equation}
under the distribution flow $(\post{0:n})_{n \in \nset}$ using \emph{sequential Monte Carlo} (SMC) methods.  The problem of computing expectations of functionals of type \eqref{eq:add:func} under these distributions will be referred to as `additive smoothing'. General state-space HMMs are prevalent in time-series and sequential-data analysis and are used extensively in, \eg, movement ecology \cite{michelot2016movehmm}, energy-consumption modeling \cite{candanedo2017methodology}, genomics \cite{yau2011bayesian}, target tracking \cite{sarkka2007rao}, enhancement and segmentation of speech and audio signals \cite{rabiner1989tutorial}; see also \cite{sarkka2013bayesian, zucchini2017hidden} and the numerous references therein. Operating on models of this sort, online additive smoothing is instrumental for, \emph{e.g.},    
\begin{enumerate}
    \item[--] \emph{path reconstruction}, \ie, the estimation of hidden states given observations. Especially in \emph{fixed-point smoothing}, where interest is in computing the expectations of $h(X_m)$ conditionally on $Y_0, \ldots, Y_n$ for some given $m$ and test function $h$ as $n$ tends to infinity, a problem that can be cast into our framework by letting, in \eqref{eq:add:func}, $\addf{m}(x_m, x_{m + 1}) = h(x_m)$ and $\addf{\ell} \equiv 0$ for all $\ell \neq m$. 
    \item[--] \emph{parameter inference}, where additive smoothing is a key ingredient in the computation of log-likelihood gradients (\emph{score functions}) via Fisher's identity or the intermediate quantity of the \emph{expectation-maximisation} (EM) \emph{algorithm}; see, \eg, \cite[Chapter~10]{Cappe:2005:IHM:1088883}. On-the-fly computation becomes especially important in online implementations via, \eg, the \emph{online EM} or \emph{recursive maximum likelihood} approaches \cite{cappe:2009,legland:mevel:1997}.  
 \end{enumerate}
